---
title: "Implementing optimized color palettes"
output: html_document
---

```{r packages}
library(tidyverse)
library(here)
library(sf) # for spatial data files
library(scales) # for improved labels
```

# Part 1: Implementing color palettes

## Import birth data

```{r}
births <- read_rds(here("data", "births.Rds"))
births
```

## Create a bar plot of average births by weekday

```{r}
total_births_weekday <- births %>%
  group_by(day_of_week) %>%
  summarize(total = sum(births))

ggplot(
  data = total_births_weekday,
  mapping = aes(x = day_of_week, y = total, fill = day_of_week)
) +
  geom_col() +
  # Turn off the fill legend because it's redundant
  guides(fill = "none")
```

#### Use a sequential palette from Color Brewer

```{r}

```

#### Highlight the weekend bars

Choose an alternative to the default palette

```{r}
total_births_weekday %>%
  mutate(weekend = day_of_week %in% c("Saturday", "Sunday")) %>%
  ggplot(_____) +
  ...
```

## Create a heatmap showing average number of births by day of year

```{r}
avg_births_month_day <- births %>%
  group_by(month, date_of_month_categorical) %>%
  summarize(avg_births = mean(births))
avg_births_month_day

birth_days_plot <- ggplot(
  data = avg_births_month_day,
  # By default, the y-axis will have December at the top, so use fct_rev() to reverse it
  mapping = aes(x = date_of_month_categorical, y = fct_rev(month), fill = avg_births)
) +
  geom_tile() +
  # Add nice labels
  labs(
    x = "Day of the month", y = NULL,
    title = "Average births per day",
    subtitle = "1994-2014",
    fill = "Average births"
  ) +
  # Force all the tiles to have equal widths and heights
  coord_equal() +
  # Use a cleaner theme
  theme_minimal()
birth_days_plot
```

#### Modify the plot to use different color palettes

```{r}
# viridis - default palette
birth_days_plot + scale_fill_viridis_c(option = "viridis")

# viridis - magma palette
birth_days_plot + scale_fill_viridis_c(option = "magma")

# color brewer - orange and red palette
birth_days_plot + scale_fill_distiller(palette = "OrRd")

# color brewer - yellow, green, and blue palette
birth_days_plot + scale_fill_distiller(palette = "YlGnBu")
```

## Import information on median household income in 2017 for Cook County, IL at the tract-level using the ACS

```{r}
cook_county_path <- here("data", "cook-county-inc.geojson")
cook_inc <- st_read(dsn = cook_county_path)
cook_inc
```

## Draw a choropleth using the median household income data

```{r}
cook_plot <- ggplot(data = cook_inc) +
  geom_sf(mapping = aes(fill = estimate, color = estimate), size = .3) +
  labs(
    fill = "Median household\nincome",
    color = "Median household\nincome"
  ) +
  theme_void()
cook_plot
```

#### Use a continuous color brewer palette to identify each tract's median household income

- Hint - use the `aesthetics` argument to duplicate a scale for multiple channels

```{r}

```

#### Use a continuous `viridis` palette to identify each tract's median household income

```{r}

```

# Part 2: Diverging color palettes

## Modify the birth day heatmap to use a diverging color palette

```{r}

```

## Modify the Cook County household income map to use a diverging color palette

```{r}

```

#### Manually specify the midpoint to be national median household income

- The median household income in 2019 was [$68,703](https://www.census.gov/library/publications/2020/demo/p60-270.html)

```{r}
# identify minimum, midrange, and maximum values
cook_inc_range <- range(cook_inc$estimate, na.rm = TRUE)
cook_inc_scale <- rescale(c(cook_inc_range[[1]], 68703, cook_inc_range[[2]]))

cook_plot +
  scale_fill_distiller(
    palette = ______,
    values = ______,
    aesthetics = c("fill", "color"),
    labels = dollar
  )
```

# Part 3: Interpolating continuous color palettes

## Draw the same choropleth for Cook County, but convert median household income into quintiles using `viridis`

#### Calculate exact quintiles

```{r}

```

#### Calculate approximate quintiles (i.e. let `viridis` choose an aesthetically pleasing set close to quintiles)

```{r}

```

#### Reverse the direction of the previous plot so that light colors are associated with lower household incomes

How does this change the emphasis of the graph?

```{r}

```

## Draw the same choropleth for Cook County median household income, but manually cut into eight groups and visualize using a diverging color brewer palette

```{r}

```

## Session Info

```{r}
sessioninfo::session_info()
```
